---
title: "Assignment1 - Overview Data"
author: "Ramón Roales-Welsch"
date: "09 January 2018"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, messages = FALSE)
```

```{r ,include=FALSE, cache=TRUE}
library(knitr)
library(readr)
library(dplyr)
library(tidytext)
library(tokenizers)
library(ggplot2)
library(tidyr)
library(scales)
library(stringr)
library(grid)
library(gridExtra)

source('create_example.R')
read_files()
gc()

```

```{r ,include=FALSE}

##Blogs
lines_df_blogs <- length(df_blogs)
df_blogs_1w <- tokenize_ngrams(df_blogs, n=1, simplify=TRUE)
df_blogs_2w <- unlist(tokenize_ngrams(df_blogs, n=2, simplify=TRUE))
df_blogs_3w <- unlist(tokenize_ngrams(df_blogs, n=3, simplify=TRUE))
words_per_line_blogs <- integer(length(df_blogs_1w))
for(i in seq(length(df_blogs_1w))){
        words_per_line_blogs[i] <- length(df_blogs_1w[[i]])
}
df_blogs_1w <- unlist(df_blogs_1w)
words_df_blogs <- length(df_blogs_1w)
word_length_blogs <- mean(nchar(df_blogs_1w))
Blogs <- c(lines_df_blogs, words_df_blogs, mean(words_per_line_blogs), word_length_blogs)
rm(df_blogs, lines_df_blogs, words_df_blogs, words_per_line_blogs, word_length_blogs)

##News
lines_df_news <- length(df_news)
df_news_1w <- tokenize_ngrams(df_news, n=1, simplify=TRUE)
df_news_2w <- unlist(tokenize_ngrams(df_news, n=2, simplify=TRUE))
df_news_3w <- unlist(tokenize_ngrams(df_news, n=3, simplify=TRUE))
words_per_line_news <- integer(length(df_news_1w))
for(i in seq(length(df_news_1w))){
        words_per_line_news[i] <- length(df_news_1w[[i]])
}
df_news_1w <- unlist(df_news_1w)
words_df_news <- length(df_news_1w)
word_length_news <- mean(nchar(df_news_1w))
News <- c(lines_df_news,words_df_news,mean(words_per_line_news),word_length_news)
rm(df_news, lines_df_news,words_df_news,words_per_line_news,word_length_news)

##Twitter
lines_df_twitter <- length(df_twitter)
df_twitter_1w <- tokenize_ngrams(df_twitter, n=1, simplify=TRUE)
df_twitter_2w <- unlist(tokenize_ngrams(df_twitter, n=2, simplify=TRUE))
df_twitter_3w <- unlist(tokenize_ngrams(df_twitter, n=3, simplify=TRUE))
words_per_line_twitter <- integer(length(df_twitter_1w))
for(i in seq(length(df_twitter_1w))){
        words_per_line_twitter[i] <- length(df_twitter_1w[[i]])
}
df_twitter_1w <- unlist(df_twitter_1w)
words_df_twitter <- length(df_twitter_1w)
word_length_twitter <- mean(nchar(df_twitter_1w))
Twitter <- c(lines_df_twitter,words_df_twitter,mean(words_per_line_twitter),word_length_twitter)
rm(df_twitter, lines_df_twitter,words_df_twitter,words_per_line_twitter,word_length_twitter)

##Overview Table
overview <- data.frame(rbind(Blogs, News, Twitter))
row.names(overview) <- c("Blogs", "News", "Twitter")

##Frequency Table
frequency <- bind_rows(mutate(data.frame(word = df_blogs_1w), author = "Blogs"),
                       mutate(data.frame(word = df_news_1w), author = "News"), 
                       mutate(data.frame(word = df_twitter_1w), author = "Twitter")) %>% 
        mutate(word = str_extract(word,"[a-z']+")) %>%
        count(author, word) %>%
        group_by(author) %>%
        mutate(proportion = n / sum(n)) %>% 
        select(-n) %>% 
        spread(author, proportion) %>% 
        gather(author, proportion, `Blogs`:`News`)

gc()
```

# Executive Summary

This is the first assignment for the Data Science Specialization Captstone Project. The project consist of a Machine Learning Application for Word Suggestions (as you have them on your phone). This assignment presents the underlying datasets from online sources:

* en_US.blogs.txt
* en_US.news.txt
* en_US.twitter.txt

The datasets are provided for the Capstone Project and can be retrieved from [Coursera](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip  "Coursera Swiftkey Datasample").

# Data Summary

```{r echo=FALSE, results='asis'}
kable(overview, caption = "Overview Table for the three Datasets")

```

# Data Plots

My first attempts were with the package 'tidytext' following instructions from https://www.tidytextmining.com/.  Nonetheless the implementation of bigrams and trigrams seemed much easier with the 'tokenizer' package. Therefore I present the results from the 'tokenizer' package.

## Results from counting single words (With Code Example)


```
# Code example

##Tokenizing

df_twitter <- readLines('source')
df_twitter_1w <- unlist(tokenize_ngrams(df_twitter, n=1, simplify=TRUE))

##Visualization

data.frame(word=df_twitter_1w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Words on Twitter (tokenizer)')

```

```{r , echo=FALSE, message=FALSE, figure=TRUE, fig.align = "center", fig.width = 6, fig.asp =  1.2}

p1 <- data.frame(word=df_blogs_1w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Words in Blogs (tokenizer)')

p2 <- data.frame(word=df_news_1w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Words in News (tokenizer)')

p3 <- data.frame(word=df_twitter_1w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Words on Twitter (tokenizer)')

grid.arrange(p1, p2, p3)
```


## bigrams (ngrams with n=2) using 'tokenizer'


```{r , echo=FALSE, message=FALSE, figure=TRUE, fig.align = "center", fig.width = 6, fig.asp =  1.2}
p1 <- data.frame(word=df_blogs_2w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Bigrams in Blogs')

p2 <- data.frame(word=df_news_2w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Bigrams in News')

p3 <- data.frame(word=df_twitter_2w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Bigrams on Twitter')

grid.arrange(p1, p2, p3)
```

## trigrams (ngrams with n=3) using 'tokenizer'

```{r , echo=FALSE, message=FALSE, figure=TRUE, fig.align = "center", fig.width = 6, fig.asp =  1.2}
p1 <- data.frame(word=df_blogs_3w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Trigrams in Blogs')

p2 <- data.frame(word=df_news_3w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Trigrams in News')

p3 <- data.frame(word=df_twitter_3w) %>%
        anti_join(stop_words) %>%
        count(word, sort = TRUE) %>%
        head(10) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
        geom_col() +
        xlab(NULL) +
        coord_flip() +
        ggtitle('Top 10 Trigrams on Twitter')

grid.arrange(p1, p2, p3)
```

## Relationship of words between datasets (With Code)

```
# Code Chunk

##Frequency Table

frequency <- bind_rows(mutate(data.frame(word = df_blogs_1w), author = "Blogs"),
                       mutate(data.frame(word = df_news_1w), author = "News"), 
                       mutate(data.frame(word = df_twitter_1w), author = "Twitter")) %>% 
        mutate(word = str_extract(word,"[a-z']+")) %>%
        count(author, word) %>%
        group_by(author) %>%
        mutate(proportion = n / sum(n)) %>% 
        select(-n) %>% 
        spread(author, proportion) %>% 
        gather(author, proportion, `Blogs`:`News`)
        
##Visualization

ggplot(frequency, aes(x = proportion, y = `Twitter`, color = abs(`Twitter` - proportion))) +
        geom_abline(color = "gray40", lty = 2) +
        geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
        geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
        scale_x_log10(labels = percent_format()) +
        scale_y_log10(labels = percent_format()) +
        scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
        facet_wrap(~author, ncol = 2) +
        theme(legend.position="none") +
        labs(y = "Twitter", x = NULL)

```

```{r , echo=FALSE, message=FALSE, figure=TRUE, fig.align = "center", fig.width = 5, fig.asp = .62}
ggplot(frequency, aes(x = proportion, y = `Twitter`, color = abs(`Twitter` - proportion))) +
        geom_abline(color = "gray40", lty = 2) +
        geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
        geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
        scale_x_log10(labels = percent_format()) +
        scale_y_log10(labels = percent_format()) +
        scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
        facet_wrap(~author, ncol = 2) +
        theme(legend.position="none") +
        labs(y = "Twitter", x = NULL)
```

# How to solve the Assignment

1. https://www.tidytextmining.com/ provides an useful approach for how to solve the assignment
    * Frequency Analysis is used
    * (Optional) Sentiment Analysis seems like a good candidate to enhance my approach
2. I will test, if a similar approach is also applicable with the 'tokenizer' package as it provided a simple and strong tool for the first visualizations.
3. The prediction system I intent to build up shall differentiate between source
    * The three datasets give three different source
    * (Optional) My shiny application shall differentiate between these source as to give the user to build his own library
    * (Optional) A fourth option will added for all datasets at the same time
    * (Optional) Inputs shall be gathered and used to individualize the predictions to the user.
        
        
*I do not know, how applicable each step is. These are all ideas, I would try to implement. If I run short of time, I will focus on the most important part, which is to create a working predictive machine.*
